---
title: "Recreating VLOOKUP in R"
author: "Matt Farrow"
date: "2019-11-13"
output: html_document
categories:
  - R
tags:
  - excel
  - dplyr
slug: recreating-vlookup-in-r
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
draft: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I recently gave a [talk on Excel](../../talk/2019-apra-excel/) that focused on pivot tables and the VLOOKUP function. Recreating the power of VLOOKUP is easy in R using the [dplyr](https://dplyr.tidyverse.org) package from the [tidyverse](https://www.tidyverse.org). 

```{r libraries, message=FALSE}
# Load dplyr package
library(dplyr)
```

One of the examples I used was collapsing research ratings into more broad "buckets" for principal, major, and leadership gift-rated prospects. We'll start be making a rating lookup table that contains all our ratings along with a corresponding "bucket." This could also be stored in a .csv or .xlsx file and loaded in separately. 

```{r lookup}
rating_lookup <-
  as_tibble(list(
    rating = c(
      "A - $2.5M to $5M",
      "B - $1M to $2.5M",
      "C - $500,000 to $1M",
      "D - $250,000 to $499,999",
      "E - $100,000 to $249,999",
      "G - $50,000 to $99,999",
      "H - $25,000 to $49,999",
      "I - $10,000 to $24,999"
    ),
    rating_bucket = c("PG", "PG", "MG", "MG", "MG", "LG", "LG", "LG")
  ))

# view our new lookup table
rating_lookup
```

Now, we'll make a sample dataset with ID numbers along with ratings.

```{r data}
df <-
  as_tibble(list(
    # create an id column using the numbers 1 through 50
    lookup_id = 1:50,
    # create a rating column by randomly sampling our previously generated ratings 50 times
    rating = sample(rating_lookup$rating, size = 50, replace = TRUE)
  ))

# look at only the first 10 observations
head(df, n = 10)
```

Now that we have our dataset as well as our lookup table, we can merge our rating buckets into our dataset using the [`left_join`](https://dplyr.tidyverse.org/reference/join.html) function.

```{r merge}
df <- left_join(df, rating_lookup, by = "rating")

# look at only the first 10 observations
head(df, n = 10)
```

Finally, we can use dplyr's [`count`](https://dplyr.tidyverse.org/reference/tally.html) function to recreate some of the power of Excel's pivot tables.

```{r pivot}
count(df, rating_bucket, name = "prospects")
```

